Scientific Reports Title to see here (20 words or less)

### 2016-09-14 draft1.tex
added section on global sensitivity analysis

$k_2$ and $\beta$ were held constant as to reduce the search space, as done by Olufsen and Ottesen. $\alpha$, the parameter used to smooth the pressure data was held constant at 1.5 to smooth the data with a minimal phase shift. The maximum firing rate, $M$, was held constant, as it only appears in this system of equations by normalizing $n$. The sympathetic nervous system delay, $\tau_d$, was held constant, owing to the propagation of discontinuities. \cite{baker1997pitfalls}

@@

\subsection*{Sensitivity Using Finite Differences}
The derivatives of all parameters were estimated using central differences. 
\begin{equation}
\frac{dh}{d\theta_j} = \frac{h(\theta_0+\frac{e_j}{2})-h(\theta_0-\frac{e_j}{2})}{\delta}
\end{equation}
where $\delta=10^{-8}*\theta_j$ and $e_j$ is a vector of length $\delta$ in the $j^{th}$ direction.
To collapse the time dimension, we calculated overall state sensitivity coefficients. \cite{stelling2004robustness}
\begin{equation}
S_{0j}(t) = \frac{1}{n_s}p_j \left(\sum_{k=1}^{n_t} \sum_{i=1}^{n_s} \left[ \frac{1}{x_i} \frac{dx_i(t_k)}{dp_j} \right]^2 \right)^{1/2}
\end{equation}
where $n_s$ = 1, as $h$ is the only state variable and $n_t$ is the number of time points available for that patient. We calculated sensitivites of parameters that were optimized and not that of those held constant.
\subsubsection*{Single Objective Sensitivity}
From the original values provided by Olufsen and Ottesen, we found that $h_0$ was the most sensitive parameter, followed by $N$, $m_{ach}$, and $m_{nor}$ in that order, of the parameters estimated. Of the four most sensitive equations, three of them ($h_0$, $m_{ach}$ and $m_{nor}$) appear directly in the equation for heart rate, and $N$ indirectly appears, as both $c_{ach}$ and $c_{nor}$ are functions of $N$.
The four estimated parameters that were the most sensitive in our study were among the five most sensitive parameters found by Olufsen and Ottesen, as seen in \ref{tab:OSSCsoriginalParams}. However, $k_2$ showed a very high degree of sensitivity in both the original and estimated parameter set, even more so than any of the estimated parameters. It also had the highest varriance in sensitivity, due to the fact that some patients exhibited a sensitivity to $k_2$ that was an order of magnitude larger than the mean.

@@

If we were to compare to the parameters estimated by Nelder Mead, we find similar sensitivites in $h_0$ and $N$ as well as in $m_{ach}$ and $m_{nor}$. The sensitivities in all of the $\tau$s included in the model is much smaller in the multiobjective case than in the single objective case. Part of this difference in sensitivity may arise from the differences in value found for these parameters by the different methods of optimization.
\section*{Discussion}
We estimated three sets of parameters that can be used to predict heart rates of patients in intensive care units. The accuracy of the model was much improved when patients were grouped by their characterstics. This improvement in accuracy is likely mostly based on that the resting heart rate of the patients in clusters more accurately matched $h_0$, the resting heart rate in the model, since $h_0$ was shown to be the most sensitive parameter in the model.

>>>>

$k_2$ and $\beta$ were held constant as to reduce the search space, as done by Olufsen and Ottesen. Additionally, when they were allowed to vary during the optimization, the parameter set generated had very small values for $m_{ach}$ and $m_{nor}$, resulting in a predicted heart rate that was basically constant at $h_0$. While this set of the parameters did mimimize the MSE, it did not capture the intrinsic varation in heart rate. $\alpha$, the parameter used to smooth the pressure data was held constant at 1.5 to smooth the data with a minimal phase shift. The maximum firing rate, $M$, was held constant, as it only appears in this system of equations by normalizing $n$. The sympathetic nervous system delay, $\tau_d$, was held constant, owing to the propagation of discontinuities. \cite{baker1997pitfalls}

@@

\subsection*{Local Sensitivity Using Finite Differences}
The derivatives of all parameters were estimated using central differences. 
\begin{equation}
\frac{dh}{d\theta_j} = \frac{h(\theta_0+\frac{e_j}{2})-h(\theta_0-\frac{e_j}{2})}{\delta}
\end{equation}
where $\delta=10^{-8}*\theta_j$ and $e_j$ is a vector of length $\delta$ in the $j^{th}$ direction.
To collapse the time dimension, we calculated overall state sensitivity coefficients. \cite{stelling2004robustness}
\begin{equation}
S_{0j}(t) = \frac{1}{n_s}p_j \left(\sum_{k=1}^{n_t} \sum_{i=1}^{n_s} \left[ \frac{1}{x_i} \frac{dx_i(t_k)}{dp_j} \right]^2 \right)^{1/2}
\end{equation}
where $n_s$ = 1, as $h$ is the only state variable and $n_t$ is the number of time points available for that patient. We calculated sensitivites of parameters that were optimized and not that of those held constant.
\subsubsection*{Single Objective Sensitivity}
From the original values provided by Olufsen and Ottesen, we found that $h_0$ was the most sensitive parameter, followed by $N$, $m_{ach}$, and $m_{nor}$ in that order, of the parameters estimated. Of the four most sensitive equations, three of them ($h_0$, $m_{ach}$ and $m_{nor}$) appear directly in the equation for heart rate, and $N$ indirectly appears, as both $c_{ach}$ and $c_{nor}$ are functions of $N$.
The four estimated parameters that were the most sensitive in our study were among the five most sensitive parameters found by Olufsen and Ottesen, as seen in \ref{tab:OSSCsoriginalParams}. However, $k_2$ showed a very high degree of sensitivity in both the original and estimated parameter set, even more so than any of the estimated parameters. It also had the highest varriance in sensitivity, due to the fact that some patients exhibited a sensitivity to $k_2$ that was an order of magnitude larger than the mean.

@@

If we were to compare to the parameters estimated by Nelder Mead, we find similar sensitivites in $h_0$ and $N$ as well as in $m_{ach}$ and $m_{nor}$. The sensitivities in all of the $\tau$s included in the model is much smaller in the multiobjective case than in the single objective case. Part of this difference in sensitivity may arise from the differences in value found for these parameters by the different methods of optimization.

\subsection{Global Sensitivity}
We used the Morris method of sampling to generate a family of parameter sets, ranging from $\pm$25\% of the nominal parameter values, with the exception of $h_0$, which was bounded between [.8, 2.0], corresponding to a resting heart rate between 48 and 120 bpm.\cite{morris1991factorial} We found that this range of parameters resulted in physiologically acceptable heart rates. As heart rates are a function of time, it was neccessary for us to collapse the time dimension to perform sensitivity analysis. We used fPCA (functional principal component analysis) to convert the heart rate curves into weights ($\omega_i$) and eigenfunctions $(\xi_i(t))$ so that $h_j(t)=\sum_i \omega_i \xi_i(t)$, and then performed sensitivity analysis on the weights. \cite{sumner2012methodology}

\section*{Discussion}
We estimated three sets of parameters that can be used to predict heart rates of patients in intensive care units. The accuracy of the model was much improved when patients were grouped by their characterstics. This improvement in accuracy is likely mostly based on that the resting heart rate of the patients in clusters more accurately matched $h_0$, the resting heart rate in the model, since $h_0$ was shown to be the most sensitive parameter in the model.
